Sperm should evolve to make female meiosis fair.

### 2014-11-17 writeup_and_notes/evolution/response_drive_MS.tex

I enjoyed reading this interesting paper. The question is clearly one where intuition fails because of the complexity of the interacting factors and the authors? models clarify these interactions.
\\
\\
\\
{\bf{R1 comment 1:}}
Models 1-3 cover the situation in which drive in female meiosis depends on sperm (padumnal) genotype or male (paternal) genotype. The models find that the invasion of such a system of drive is unlikely because in a random-mating population drive is very rare because both sexual partners need to carry the rare allele and because drive, when it does occur, creates driving homozygotes which are assumed to have low viability and which cannot themselves undergo drive. If there is no cost to homozygotes the ?driver? should go to fixation slowly at first. Forms of inbreeding would probably also facilitate initial invasion.

{\bf{Response to R1 comment 1:}} \emph{We have added inbreeding to our analyses of models 1-2, see response to AE comment 1}
\\
\\
\\
{\bf{R1 comment 2:}}
Models 4-5 cover the situation in which an allele drives in female meiosis and this is modified by sperm genotype at the same locus (by assumption of tight linkage). Model 6 considers an unlinked sperm modifier. The interesting case in Model 4 is the suppressor allele B-. This allele initially does not suppress its own drive (because rare) but suppresses the drive of B and thus ends up in A/B- heterozygotes rather than B/B- heterozygotes. The authors say ``avoid the excess formation of low fitness homozygotes'' but I think this is slightly misleading because B and B- are different ?alleles?. B- displaces B and often comes to a lower equilibrium frequency presumably because of the effects of suppressing its own drive
(B- causes the production of higher fitness A/B- heterozygotes rather than lower fitness B-/B- homozygotes. If the difference in fitness is not too great then B- ?loses? from suppressing its own drive).

{\bf{Response to R1 comment 2:}} \emph{We have changed this section to (e.g. changing ``avoid the excess formation of low fitness homozygotes'' to ``avoid the excess formation of low fitness genotypes'') in accordance with these suggestions.}
\\
\\
\\
{\bf{R1 minor comments:}}
\\ line 272 ``so it avoid[s]'' line 274 ``the the''
\\
The take home message is that sperm alleles usually evolve to suppress female meiotic drive, whether linked (in coupling or repulsion) or unlinked to the driving locus.
\\
line 339 ``novel females meiotic drivers'' rather ``novel female meiotic drivers'' or ``novel females' meiotic drivers''
\\
line 342 ``Tthese''

{\bf{Response to R1 minor comments:}} \emph{We fixed these, thanks!}

@@

Although I found the motivating question and background biology really interesting, I am not really convinced by the analysis.

{\bf{Response to R2 summary:}} \emph{We appreciate that you find the question, the biology, and the results compelling, and are disappointed that you are unconvinced by the analysis. However, we can find no statement in your specific comments below that suggest that you are unconvinced. We therefore respond to the comments below, and hope that the responses to your specific comments are convincing.}
\\
\\
\\
Some comments:
\\
{\bf{R2 comment 1 :}}
(1) The results are pretty obvious a priori, and don?t really warrant the development and analysis of explicit models. As the authors note in the Introduction, the matching-alleles models are instances of greenbeard genes, and we already know that these are under positively frequency-dependent selection, and cannot invade from rarity in unstructured populations if there is any obligate cost involved, such as the cost of the machinery that does the allelic recognition. (This link also needs to be made in the Results and Discussion sections.) And in the scenarios where sperm are deciding the fate of drivers in an egg to which they have no genetic relatedness, it is obvious that their only concern will be to maximize the viability of the ensuing zygote, which means suppressing drive.

{\bf{Response to R2 comment 1:}} \emph{We believe that intuition without rigorous theoretical support can be dangerous. Therefore, while the results may be intuitive to R2, they were not to ourselves or to R1. We therefore believe that our results are not trivial and represent an important contribution. The complexity of our analysis [and the bit that can choke simple intuition] is that a model of drive that maintains polymorphism requires negative frequency dependnce [e.g. in the form or of recessivity], and that the benefit of the geenbeard is [as R2 points out] positively frequency depednent. This makes co-existance [i.e. a joint polymorphism] unlikely. We note that the inutition of R2 with regard to our analysis appears somewhat off-base -- under some non-zero costs to a self-promoting driver, it can sweep through a population [Figure 1], a simple result that R2 does not explain with his simple intuitive assertions, above.}
\\
\\
\\
{\bf{R2 comment 2:}} (2) Explicit models would be a lot more useful if the scope of the study were extended to inbred populations, where there are more selective pressures at play and where it is less obvious to see what will be favoured. Here, driving genes would be a lot more likely to encounter identical-by-descent copies in sperm, which could act to promote collaboration on the part of the sperm (analogous to the boost that greenbeards get in structured populations, owing to increased rate of encounters). But drivers would also suffer from the reduced frequency of heterozygotes under inbreeding, and this would tend to inhibit drive. And there may be additional pressures owing to inbreeding favouring bias sex ratios, which driving sex chromosomes or, indeed, whole genomes could be important in bringing about, under female heterogamety e.g. see Gardner \& Ross (2014, ?Mating ecology explains patterns of genome elimination?, Ecol Lett, doi: 10.1111/ele.12383). An explicit model would be doing really useful work here, showing which of these effects is more important, and why. The present results for the outbreeding case, that seem a bit too trivial to stand on their own, would usefully serve to prepare the reader for more substantial results for inbreeding scenarios.

### 2014-11-17 writeup_and_notes/evolution/RESUBMISSIONmale_dependent_drive.tex
minor changes to response letter. acknowledgements added to ms... graham add more

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
\date{}
\begin{document}
\maketitle
\begin{center}
$^1$Department of Plant Biology, \\ University of Minnesota --€“ Twin
Cities. St. Paul MN, 55108 \\
$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
\end{center}

@@

  \appendix

@@

\bibliography{refs}

>>>>

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0.0pt}
\rfoot{}
\cfoot{\thepage}
\date{}
\begin{document}
\maketitle
\begin{center}
$^1$Department of Plant Biology, \\ University of Minnesota -- Twin
Cities. St. Paul MN, 55108 \\
$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
\end{center}

@@

\newpage 
\section*{Acknowledgements}
We would like to thank members of the Coop lab for comments. This work was made possible by the NSF Postdoctoral Fellowship in Biology FY 2010 project number 1002942 awarded to Y.B..
\newpage 
\bibliography{refs}
\newpage 
  \appendix

@@
